The variability of Betelgeuse explained by surface convection\thanks{Based on observations obtained at the T\'elescope Bernard Lyot (TBL) at Observatoire du Pic du Midi, CNRS/INSU and Universit\'e de Toulouse, France.}

### 2024-05-17 main.tex
test 6
test 7
fix typos

\section{Introduction}
Betelgeuse is a prototypical red supergiant (RSG), known to be a semi-regular variable. Several periods can be found in the 
literature, usually clustered around the so-called long secondary period (LSP) of 2000 d, plus shorter periods around 400 and 200 d \citep{kiss_variability_2006}. The origin of the LSP remains a source of discussion: some have tried to find the explanation of the LSP in the lifetime of the large convective cells present at the surface of RSG \citep[e.g][]{stothers_giant_2010} while others have invoked a magnetic field \citep{wood_long_2004}. The shorter periods of 400 and 200 d are given a different origin than the LSP. By studying a large sample of RSGs, \cite{kiss_variability_2006} attributed the 400 d period to radial pulsation modes, while the 200 d period being attributed to the first overtone \citep{joyce_standing_2020}, in agreement with the historical work of \cite{stothers_pulsation_1969}.\\

Besides radial pulsation, surface activity leading to random brightness variations has also been mentioned to explain the variability of Betelgeuse, firstly theorised by \cite{schwarzschild_scale_1975}. More recently, \cite{gray_mass_2008} proposed that the 400 d period was a consequence of convective activity. Betelgeuse is known to present large convective cells with lifetimes of the order of one to two years \citep{lopez_ariste_convective_2018} and measurable changes 
within the span of one week. Bright convection cells near the disk center will increase the integrated brightness of the star compared with other situations where such cells are found near the edges. So, even without invoking the formation of 
dust or other changes in opacity, one may argue that the simple evolution of convective patterns of the star may create variability. Such 
variability could be expected to be random, but will present quasi-periodicities related to the typical time scales of the 
convective patterns \citep{gray_mass_2008}. \\

@@

Betelgeuse has been observed for the last 10 years with Narval and Neo-Narval at the Telescope Bernard Lyot\footnote[1]{https://tbl.omp.eu/}. These two instruments measure the polarization 
over the visible and near infrared spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. The signal-to-noise ratios are not sufficient to measure the weak polarization signals in individual atomic lines of the 
spectrum. These amplitudes are known to be of the order of $10^{-4}$ times the continuum intensity, and they only exceptionally reach amplitudes of $10^{-3}$. When these high amplitudes are present, enough photons can be accumulated per spectral bin to 
detect the linear polarization signal above noise in lines \citep{auriere_discovery_2016}. Most commonly, the amplitudes of linear polarization are below noise levels. On those 
occasions, we have to add up the signals of thousands of lines to increase the signal-to-noise ratios. This addition 
is performed through a technique called Least-Squares Deconvolution (LSD) \citep{donati_spectropolarimetric_1997}.% and assumes that the 
%linear polarization signal is similar 
%in all spectral lines up to scale factors in the sampling and amplitude.
This technique has been successfully used in the past to measure 
magnetic field distributions over stellar surfaces and is now employed to produce images of the brightness distributions in the photosphere 
of Betelgeuse and other RSGs, as cited in the Introduction.

@@

We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, ranging from -25 km/s to +60 km/s to cover the entire 
signal present in the LSD profile. \cite{lopez_ariste_convective_2018} identified the most blueshifted signal towards -20 km/s and interpreted it as  the maximum velocity of the rising plasma. The most redshifted signal was often found at +40 km/s, interpreted 
as the rest velocity of the star. Sometimes, signals at velocities greater than +40 km/s can be found, which can be interpreted either as dark and cold plasma falling back to Betelgeuse or as rising plumes of plasma located on the hidden face of Betelgeuse, ascending so high that they appear above the limb \citep[for the case of the RSG $\mu$Cep]{lopez_ariste_height_2023}.

@@

Before summing up the results of our work, we can afford to propose an explanation for the variability of Betelgeuse after the great dimming with surface convection. RSGs experience mass loss events when plasma is ejected into the interstellar medium \citep{josselin_atmospheric_2007}. 
These events have been inferred in RSG such as $\mu$Cep at photospheric heights, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star, attributing it to rising plumes of plasma in the back hemisphere of the star. 
In the case of Betelgeuse, the excess of linear polarization beyond the limb is rare and not as pronounced as those observed in $\mu$Cep. They do not appear to contribute to 
 the variability of the star. However, since the great dimming, the variability of Betelgeuse has changed. This suggests that the great dimming differed significantly from the usual mass loss event, and perhaps it hustled the  photosphere. \

>>>>

\section{Introduction}
\label{sec:introduction}
Betelgeuse is a prototypical red supergiant (RSG), known to be a semi-regular variable. Several periods can be found in the 
literature, usually clustered around the so-called long secondary period (LSP) of 2000 d, plus shorter periods around 400 and 200 d \citep{kiss_variability_2006}. The origin of the LSP remains a source of discussion: some have tried to find the explanation of the LSP in the lifetime of the large convective cells present at the surface of RSGs \citep[e.g][]{stothers_giant_2010} while others have invoked a magnetic field \citep{wood_long_2004}. The shorter periods of 400 and 200 d are given a different origin than the LSP. By studying a large sample of RSGs, \cite{kiss_variability_2006} attributed the 400 d period to radial pulsation modes, while the 200 d period being attributed to the first overtone \citep{joyce_standing_2020}, in agreement with the historical work of \cite{stothers_pulsation_1969}.\\

Besides radial pulsation, surface activity leading to random brightness variations has also been mentioned to explain the variability of Betelgeuse, first theorised by \cite{schwarzschild_scale_1975}. More recently, \cite{gray_mass_2008} proposed that the 400 d period was a consequence of convective activity. Betelgeuse is known to present large convective cells with lifetimes of the order of one to two years \citep{lopez_ariste_convective_2018} and measurable changes 
within the span of one week. Bright convection cells near the disk center will increase the integrated brightness of the star compared with other situations where such cells are found near the edges. So, even without invoking the formation of 
dust or other changes in opacity, one may argue that the simple evolution of convective patterns of the star may create variability. Such 
variability could be expected to be random, but will present quasi-periodicities related to the typical time scales of the 
convective patterns \citep{gray_mass_2008}. \\

@@

Betelgeuse has been observed for the last 10 years with Narval and Neo-Narval at the Telescope Bernard Lyot\footnote[1]{https://tbl.omp.eu/}. These two instruments measure the polarization 
over the visible and near infrared spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. The signal-to-noise ratios are not sufficient to measure the weak polarization signals in individual atomic lines of the 
spectrum. These amplitudes are known to be of the order of $10^{-4}$ times the continuum intensity, and they only exceptionally reach amplitudes of $10^{-3}$. When these high amplitudes are present, enough photons can be accumulated per spectral bin to 
detect the linear polarization signal above noise in lines \citep{auriere_discovery_2016}. Most commonly, the amplitudes of linear polarization are below noise levels. On those 
occasions, we have to add up the signals of thousands of lines to increase the signal-to-noise ratios. This addition 
is performed through a technique called Least-Squares Deconvolution \citep[LSD;][]{donati_spectropolarimetric_1997}.% and assumes that the 
%linear polarization signal is similar 
%in all spectral lines up to scale factors in the sampling and amplitude.
This technique has been successfully used in the past to measure 
magnetic field distributions over stellar surfaces and is now employed to produce images of the brightness distributions in the photosphere 
of Betelgeuse and other RSGs, as cited in section~\ref{sec:introduction}.

@@

We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, ranging from -25 km/s to +60 km/s to cover the entire 
signal present in the LSD profile. \cite{lopez_ariste_convective_2018} identified the most blueshifted signal towards -20 km/s and interpreted it as  the maximum velocity of the rising plasma. The most redshifted signal was often found at +40 km/s, interpreted 
as the rest velocity of the star. Sometimes, signals at velocities greater than +40 km/s can be found, which can be interpreted either as dark and cold plasma falling back to Betelgeuse or as rising plumes of plasma located on the hidden face of Betelgeuse, ascending so high that they appear above the limb \citep[for the case of the RSG $\mu$~Cep]{lopez_ariste_height_2023}.

@@

Before summing up the results of our work, we can afford to propose an explanation for the variability of Betelgeuse after the great dimming with surface convection. RSGs experience mass loss events when plasma is ejected into the interstellar medium \citep{josselin_atmospheric_2007}. 
These events have been inferred in RSG such as $\mu$~Cep at photospheric heights, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star, attributing it to rising plumes of plasma in the back hemisphere of the star. 
In the case of Betelgeuse, the excess of linear polarization beyond the limb is rare and not as pronounced as those observed in $\mu$~Cep. They do not appear to contribute to 
 the variability of the star. However, since the great dimming, the variability of Betelgeuse has changed. This suggests that the great dimming differed significantly from the usual mass loss event, and perhaps it hustled the  photosphere. \
